---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Saima Arina"
date: "02/15/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r - filter dennys for Alaska}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```

There are 3 Denny's locations in Alaska.

```{r - filter la quintas for Alaska}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```

There are 3 La Quinta locations in Alaska.

### Exercise 2

```{r - how many pairings between all dennys and la quinta locations in AK}
nrow(dn_ak) * nrow(lq_ak)
```

There are 6 pairings we need to calculate.

### Exercise 3

```{r - join dataframes}
dn_lq_ak <- full_join(dn_ak, lq_ak,
  by = "state"
)
dn_lq_ak
```

### Exercise 4

```{r - observations and variables}
nrow(dn_lq_ak)
ncol(dn_lq_ak)
```

There are 6 observations in the joined data frame. The names of the variables in this data frame is: address, city, state, zip, longitude, and latitude. 


### Exercise 5

We use the function "mutate" from the tidyverse to add a new variable to a data frame while keeping the existing variables. 

```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) # distance in km
}
```


### Exercise 6

```{r - calculate distance between all pairings}
dn_lq_ak <- dn_lq_ak %>% 
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```

### Exercise 7


